2.1.1 Canpath Data Check

### 2024-03-12 2_scripts/2.1.1 Canpath Data.qmd
update Canpath cleaning code

# The serology data
Canpath_ser <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Results_74503par.csv")
```

@@

-   All for three: pid (participants ID) and sid (sample ID)
```{r}
# sample_ID

@@

# Research ID
# qx data:
summary(Canpath_qx$ResearcherID)                        # N=96014, no missing
table(is.na(Canpath_qx$ResearcherID), useNA = 'ifany')
length(unique(Canpath_qx$ResearcherID))                 # all 96014 rid are unique

@@

  ## find duplicates
  n_admin <- data.frame(table(Canpath_admin$ResearcherID))
  n_admin[n_admin$Freq > 1,]          # 3 duplicates, sample pp with two samples (diff sample ID)

# ser data
summary(Canpath_ser$ResearcherID)                      # N=74503, no missing
table(is.na(Canpath_ser$ResearcherID), useNA = 'ifany')
length(unique(Canpath_ser$ResearcherID))               # N=25975, duplicates
```
#### 1.2 Merge three data
```{r}
# select the variables we need from each subset:
# qx subset:
qx_s<-select(Canpath_qx, ResearcherID, C_ADM_STUDY_DATASET, C_SDC_AGE, C_SDC_SEX,
             C_ADM_FSA)
# apply shorter var names:
colnames(qx_s)<-c("rid", "study_id", "age", "sex", "fsa")

# admin subset
admin_s<-select(Canpath_admin, ResearcherID, C1_SAMPLE_ID,
                C1_ADM_COHORT_ID, C1_ADM_COLLECT_DATE, C1_SAMPLE_TYPE, 
                C1_SAMPLE_PROCESS_DATE)
colnames(admin_s)<-c("rid", "sid", "cohort_id", "date_col", "stype", "date_pro")

# ser data, use anti-N only
ser_s<-select(Canpath_ser[Canpath_ser$C1_CITF_ASSAY_ID %in% c("10010-00", "20010-03"),],
              ResearcherID, C1_ADM_SPECIMEN_ID, C1_CITF_ASSAY_ID, C1_SAMPLE_DATE_ASSAY_RUN,
              C1_SAMPLE_RESULT, C1_SAMPLE_RESULTS_DESCRIPTION)
colnames(ser_s)<-c("rid", "sid", "assay_id", "date_asy", "result", "pos")

# reshape the subset
# ser_sw<-reshape(ser_s, idvar = "sid", timevar = "assay_id", direction = "wide")
# reshape is unnecessary
```
```{r}
# merge those small subsets
m1<-merge(admin_s, ser_s[!is.na(ser_s$rid) & !is.na(ser_s$sid),],
          by = c("rid", "sid"), all.x = F, all.y = T)
m2<-merge(m1, qx_s, by='rid', all.x = T, all.y = F)

ser_s %>%
  .[!is.na(.$rid) & is.na(.$sid),]

@@

>>>>

# The serology data
Canpath_ser <- read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Serology_Results_74503par.csv")

# The new MTP Qx subset
Canpath_qx_MTP<-read.csv("../1_data/private/CANPATH/DAO-543759_ResearcherDataset_Qx_1114par_1127varMTP.csv")
```

@@

-   All for three: pid (participants ID) and sid (sample ID)
```{r}
# sample_ID: only available in admin and ser data

@@

# Research ID: invididual ID
# qx data:
summary(Canpath_qx$ResearcherID)                        # N=96014, no missing
table(is.na(Canpath_qx$ResearcherID), useNA = 'ifany')
length(unique(Canpath_qx$ResearcherID))                 # all 96014 rid are unique

@@

  ## find duplicates
  n_admin <- data.frame(table(Canpath_admin$ResearcherID))
  n_admin[n_admin$Freq > 1,]          # 3 duplicates, one pp with two samples (diff sample ID), collected on diff day. Drop the first one

# ser data
summary(Canpath_ser$ResearcherID)                      # N=74503, no missing
table(is.na(Canpath_ser$ResearcherID), useNA = 'ifany')
length(unique(Canpath_ser$ResearcherID))               # N=25975, duplicates due to same sample tested for multiple assays
```
#### 1.2 Merge three data
```{r}
# select the variables we need from each subset:
# qx subset:
qx_s<-select(Canpath_qx, ResearcherID, C_ADM_STUDY_DATASET, C_SDC_AGE, C_SDC_SEX,
             C_ADM_FSA)

# don't forget MTP subset
qx_s_mtp<-select(Canpath_qx_MTP, ResearcherID, C_ADM_STUDY_ID, 
             C_SDC_AGE, C_SDC_SEX, C_ADM_FSA,
             C1_SDC_AGE, C1_SDC_SEX, C1_ADM_FSA)
# apply shorter var names:
colnames(qx_s)<-c("rid", "study_id", "age", "sex", "fsa")
colnames(qx_s_mtp)<-c("rid", "study_id", "age", "sex", "fsa", 
                      "age1", "sex1", "fsa1")

# identify those missed from original data:
qx_s_mtp[qx_s_mtp$rid %in% qx_s$rid,] # not missed, N=699
qx_mtpmiss<-qx_s_mtp[!(qx_s_mtp$rid %in% qx_s$rid),] # missed, N=445

# rbind those two into one:
qx_s<-rbind.data.frame(qx_s, 
                       select(qx_mtpmiss, -c("age1", "sex1", "fsa1")))

# clean study_id
qx_s$study_id[qx_s$study_id=='5']<-"MTP"
table(qx_s$study_id, useNA = 'ifany')

# admin subset
admin_s<-select(Canpath_admin, ResearcherID, C1_SAMPLE_ID,
                C1_ADM_COHORT_ID, C1_ADM_COLLECT_DATE, C1_SAMPLE_TYPE, 
                C1_SAMPLE_PROCESS_DATE)
colnames(admin_s)<-c("rid", "sid_admin", "cohort_id", "date_col", "stype", "date_pro")

# ser data, use anti-N only
ser_s<-select(Canpath_ser[Canpath_ser$C1_CITF_ASSAY_ID %in% c("10010-00", "20010-03"),],
              ResearcherID, C1_ADM_SPECIMEN_ID, C1_CITF_ASSAY_ID, C1_SAMPLE_DATE_ASSAY_RUN,
              C1_SAMPLE_RESULT, C1_SAMPLE_RESULTS_DESCRIPTION)
colnames(ser_s)<-c("rid", "sid_ser", "assay_id", "date_asy", "result", "pos")

table(ser_s$rid %in% qx_mtpmiss$rid, useNA = 'ifany')
table(admin_s$rid %in% qx_mtpmiss$rid, useNA = 'ifany')

# reshape the subset
# ser_sw<-reshape(ser_s, idvar = "sid", timevar = "assay_id", direction = "wide")
# reshape is unnecessary
```
#### Data quality check of those small dataset
```{r}
# admin_s: missing
table(is.na(admin_s$rid), useNA = 'ifany')     # rid no missing, not all unique (3 duplicates)
# remove duplicates: drop the earliest sample
admin_s<-admin_s[!admin_s$sid %in% c("DB0006218", "DB0009025", "DB0009310"),]

table(is.na(admin_s$sid), useNA = 'ifany')     # sid 8349 missing
length(unique(admin_s$sid[!is.na(admin_s$sid)])) # non-missing sids all unique

# ser_s (only 2 anti-N assays): missing
table(is.na(ser_s$rid), useNA = 'ifany')      # rid no missing
length(unique(ser_s$rid))                     # duplicates exists

  n_ser <- data.frame(table(ser_s$rid))
  n_ser[n_ser$Freq > 1,]
# some sample has been tested multiple times by the same assay
# none of those testing has been result in inconsistency of anti-N positivity

# in ser_s, see where rid is duplicated: all used Mount Sinai Hosp Assay (anti-N)
ser_rid_dup<-ser_s[ser_s$rid %in% n_ser[n_ser$Freq > 1,]$Var1, ]
table(ser_s[ser_s$rid %in% n_ser[n_ser$Freq > 1,]$Var1, ]$assay_id, useNA='ifany')

# need to deduplicate ser_s, only keep one of duplicated records
ser_rid_dup$id<-c(1:404)
ser_rid_dup %>%
  distinct(rid, .keep_all=TRUE) # N=163 （404-163 = 241 de-duplicated)

ser_s<-ser_s %>%
  distinct(rid, .keep_all=TRUE) # N=25839, number matches
```

```{r}
# merge those small subsets
m1<-merge(admin_s, ser_s,
          by = c("rid"), all.x = F, all.y = T)
m2<-merge(m1, qx_s, by='rid', all.x = T, all.y = F)

# sid match
m2$sid_match[!is.na(m2$sid_admin) & !is.na(m2$sid_ser) &
             m2$sid_admin == m2$sid_ser]<-"match"
m2$sid_match[!is.na(m2$sid_admin) & !is.na(m2$sid_ser) &
             m2$sid_admin != m2$sid_ser]<-"mismatch"
m2$sid_match[!is.na(m2$sid_admin) & is.na(m2$sid_ser)]<-"ser miss"
m2$sid_match[is.na(m2$sid_admin) & !is.na(m2$sid_ser)]<-"admin miss"
m2$sid_match[is.na(m2$sid_admin) & is.na(m2$sid_ser)]<-"all miss"
table(m2$sid_match, useNA = 'ifany')

# work on this part later #

@@

##===================== Part.II Estimate Seroprevalence =================

**Notes:** Seroprevalence estimation requires individual analysis and a little bit weighting. Those who has been sampled multiple times should be down-weighted properly. Those who are sampled multiple times in different months will not be down-weighted.

M.Knight's data cleaning strategy is much more straightforward, as he only need to locate the participants who provided sero data, so matching on researcherID makes senses. In our case we need to treat sero data more carefully, as in Canpath sero data, one individual could provide multiple samples, and those samples were tested by multiple assays (Abbott N/S, Mount Sinai Hosp N/S/S-RBD).

```{r}

```

##
